%!TEX root = paper.tex

\section{The Spack Package Manager}
\label{sec:implementation}
Based on our experiences at LLNL, we have developed
{\it Spack}, the Supercomputing PACKage manager.
Spack is written in Python.  We chose Python because it is
very widely used int he HPC community, and because it provides
flexible scripting capabilities. 
%
Like some prior systems, Spack supports arbitrarily many software installations.
It borrows ideas from Nix; packages are installed in their own prefixes,
and configurations are managed using hashes.
% 
Spack adds several unique capabilities that are essential for HPC:
\begin{enumerate}
\item {\bf Parametric builds} address the matrix problem:
      Packages are parameterized by name, version, architecture, compiler, 
      options, and dependencies.
\item To manage the configuration space, spack provides a novel, 
      recursive {\bf spec syntax} for dependency graphs.
\item Specs support versioned, ABI-incompatible interfaces like MPI through
      {\bf versioned virtual dependencies}.
\item Spack builds with {\bf compiler wrappers} that reduce the burden of build
      consistency on the package author.
\end{enumerate}


\input{implementation-packages}

\subsection{Spack Specs}\label{sec:specs}

  Specs are partial descriptions of full
dependency graphs, but they are concise.  This allows a user to use
only a few parameters to quickly identify a configuration to install,
or to query existing installations and their dependencies.



\todo{1 page}



\subsection{Versioned Virtual Dependencies}
	\todo{.5 page}

\subsection{Abstract \& Concrete Specs}
	\todo{.75 page}
	

